Portfolios of Subgraph Isomorphism Algorithms

### 2015-12-08 paper/paper.tex
restructure and add ECDF graph

\usepackage{complexity}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{cleveref}                  % no need to type Figure etc
\usepackage{xcolor,colortbl}

@@

\section{Per-Instance Algorithm Selection}

@@

\subsection{Subgraph isomorphism algorithms}

The Glasgow algorithm has a parameter, which controls the lengths of paths used when reasoning about
non-adjacent vertices.  In experiments reported in \cite{McCreesh:2015}, the choice of paths of
length 3 was used as a reasonable compromise---longer paths lead to prohibitively expensive
preprocessing on larger, denser instances. This is often not the best choice on an instance by
instance basis: sometimes path-based reasoning gives no benefit at all, sometimes considering only
paths of length 2 suffices, occasionally paths of length 4 are helpful, and even looking at paths of
length 3 is relatively expensive on some graphs. Therefore, in this new study, we consider 4
different instances of the Glasgow algorithm, with path lengths varying from 1 to 4.

We also introduce two new variants of LAD. The first variant is a weaker LAD filtering (called LAD
preprocessing) which is applied once, without performing a backtracking search, and detects
inconsistencies on many instances: For each pattern node $u$, we check that there exists at least
one target node $v$ such that for each neighbor $u'$ of $u$ there exists a different neighbor $v'$
of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$. LAD
preprocessing is an incomplete algorithm as it can only detect inconsistency of some instances: when
it does not detect inconsistency, the instance may either be consistent or not. The time complexity
of LAD preprocessing is ${\cal O}(n(n'+p'))$ where $n$ and $n'$ are the number of pattern and target
nodes, respectively, and $p$ is the number of target edges.

The second variant of LAD is called new LAD and it is stronger than the default LAD introduced in
\cite{Solnon:2010}. It combines LAD filtering with new invariant properties borrowed from SND
\cite{Audemard:2014}. The idea is to label each edge $(u,v)$ with the number of paths of length 2
between $u$ and $v$, and each node $u$ with the number of cycles of length 3 passing through $u$,
and to add the constraint that the label of a pattern node (resp. edge) must be smaller than or
equal to the label of its associated target node (resp. edge).

\subsection{Problem instances}

@@

@@

\subsection{Experimental Setup}

\subsection{Algorithm Selection approach}

@@

@@

Figure~\ref{fig:portfolio-ecdf} shows the cumulative density function for the
individual solvers, the virtual best solver, and the LLAMA portfolio. The
virtual best solver clearly dominates with a significant margin to the next
best. VF2 is far below all other solvers. The portfolio does not perform well
for instances that can be solved quickly because of the overhead incurred
through feature computation. As the instances become more difficult to solve,
its performance improves. For instances that take more than 1,000 seconds to
solve, the portfolio eclipses any individual solver and closes most of the gap
to the virtual best solver.

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figures/portfolio-ecdf}
\caption{Empirical cumulative density function of solvers, virtual best solvers,
and the LLAMA portfolio.}
\label{fig:portfolio-ecdf}
\end{figure}

>>>>

\usepackage{times}
\usepackage{complexity}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{cleveref}                  % no need to type Figure etc
\usepackage{xcolor,colortbl}

@@

\section{Background}

\subsection{Per-Instance Algorithm Selection}

@@

\subsection{Subgraph Isomorphism}

\todo[inline]{Add some background on subgraph isomorphism.}

\section{Experimental Setup}

\subsection{Problem Instances}

@@

\subsection{Subgraph Isomorphism Algorithms}

We also introduce two new variants of LAD. The first variant is a weaker LAD filtering (called LAD
preprocessing) which is applied once, without performing a backtracking search, and detects
inconsistencies on many instances: For each pattern node $u$, we check that there exists at least
one target node $v$ such that for each neighbour $u'$ of $u$ there exists a
different neighbour $v'$
of $v$ such that the degree of $u'$ is smaller than or equal to the degree of $v'$. LAD
preprocessing is an incomplete algorithm as it can only detect inconsistency of some instances: when
it does not detect inconsistency, the instance may either be consistent or not. The time complexity
of LAD preprocessing is ${\cal O}(n(n'+p'))$ where $n$ and $n'$ are the number of pattern and target
nodes, respectively, and $p$ is the number of target edges.

The second variant of LAD is called new LAD and it is stronger than the default LAD introduced in
\cite{Solnon:2010}. It combines LAD filtering with new invariant properties borrowed from SND
\cite{Audemard:2014}. The idea is to label each edge $(u,v)$ with the number of paths of length 2
between $u$ and $v$, and each node $u$ with the number of cycles of length 3 passing through $u$,
and to add the constraint that the label of a pattern node (resp.\ edge) must be smaller than or
equal to the label of its associated target node (resp.\ edge).

The Glasgow algorithm has a parameter, which controls the lengths of paths used
when reasoning about non-adjacent vertices.  In experiments reported in
\cite{McCreesh:2015}, the choice of paths of length 3 was used as a reasonable
compromise -- longer paths lead to prohibitively expensive preprocessing on
larger, denser instances. This is often not the best choice on an instance by
instance basis: sometimes path-based reasoning gives no benefit at all,
sometimes considering only paths of length 2 suffices, occasionally paths of
length 4 are helpful, and even looking at paths of length 3 is relatively
expensive on some graphs.

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figures/ecdf-glasgow}
\caption{Empirical cumulative distribution function for the Glasgow algorithm
with path lengths 1, 2, 3, and 4.}
\label{fig:ecdf-glasgow}
\end{figure}

Figure~\ref{fig:ecdf-glasgow} shows the cumulative distribution function for the
Glasgow algorithm with four different path lengths. As the instances become
harder to solve, longer paths become better. This is what we expect, as more
reasoning is expensive, but potentially increases the reduction of the search
space. In particular, the relative performance of the algorithms is monotone in
the time it takes to solve the instance -- path length 1 is best at first, but
gets worse as the time to solve the instance increases until path length 2 is
better. After that point, length 1 is never better than length 2 again. The same
is true for the other path lengths.

The figure illustrates the potential for portfolios and algorithm selection we
have. There is clearly no path length that dominates throughout. Therefore, we
consider 4 different instances of the Glasgow algorithm here, with path lengths
varying from 1 to 4.

@@

\section{Algorithm Selection Approach}

@@

\subsection{Experimental Environment}

@@

Figure~\ref{fig:portfolio-ecdf} shows the cumulative distribution function for
the individual solvers, the virtual best solver, and the LLAMA portfolio. The
virtual best solver clearly dominates with a significant margin to the next
best. VF2 is far below all other solvers. The portfolio does not perform well
for instances that can be solved quickly because of the overhead incurred
through feature computation. As the instances become more difficult to solve,
its performance improves. For instances that take more than 1,000 seconds to
solve, the portfolio eclipses any individual solver and closes most of the gap
to the virtual best solver.

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{figures/portfolio-ecdf}
\caption{Empirical cumulative distribution function of solvers, virtual best
solvers, and the LLAMA portfolio.}
\label{fig:portfolio-ecdf}
\end{figure}
